---
title: "Titanic Survival"
author: "Binod Jung Bogati"
date: "2/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Titanic: Machine Learning from Disaster

- Titanic sinking on April 15, 1912
- Data provided on our website goo.gl/DYzSmA
- 809 out of 1309 passengers got killed
- Task
  - I Can we predict who survived?
  - I Why did people die / Which groups?


### Load the packages and data

```{r}
library(mlr)

load("titanic.rda")

str(data)
```

### Pre-processing

```{r}
print(summarizeColumns(data)[, -c(5, 6, 7)], digits = 0)

# set empty factor levels to NA

data$Embarked[data$Embarked == ""] = NA
data$Embarked = droplevels(data$Embarked)
data$Cabin[data$Cabin == ""] = NA
data$Cabin = droplevels(data$Cabin)


```

```{r}
library(BBmisc)
library(stringi)

# price per person, multiple tickets by one person

data$farePp = data$Fare / (data$Parch + data$Sibsp + 1)

# the deck can be extracted from the cabin number

data$deck = as.factor(stri_sub(data$Cabin, 1, 1))

# Starboard had an odd number, portside even cabin numbers

data$portside = stri_extract_last_regex(data$Cabin, "[0-9]")
data$portside = as.numeric(data$portside) %% 2

# drop stuff we cannot easily model on

data = dropNamed(data, 
                 c("Cabin", "PassengerId", "Ticket", "Name"))

print(summarizeColumns(data)[, -c(5, 6, 7)], digits = 0)
```

- Impute missing numeric values with median, missing factor values with a separate category

- NB: This is really naive, we should probably embed this in cross-validation


```{r}
# imputation

data = impute(data, cols = list(
  Age = imputeMedian(), 
  Fare = imputeMedian(), 
  Embarked = imputeConstant("__mins__"), 
  farePp = imputeMedian(), 
  deck = imputeConstant("__mins__"), 
  portside = imputeConstant("__mins__")
))

data = data$data

data = convertDataFrameCols(data, chars.as.factor = TRUE)

head(data)
```

### Task

```{r}
# making task 
task = makeClassifTask(id = "titanic", data = data,
                       target = "Survived", positive = "1")

print(task)
```

### What MLR algorithms are available?

```{r}
tab = listLearners(task, warn.missing.packages = FALSE)
tab[1:5, c("class", "package")]
```

### Learner

```{r}
lrn = makeLearner("classif.kknn", k = 3, predict.type = "prob")

print(lrn)
```

## Train

```{r}
set.seed(123)

n = getTaskSize(task)

train = sample(n, size = 2/3 * n)

test = setdiff(1:n, train)

head(sort(train))

head(sort(test))

mod = train(lrn, task, subset = train)

```

### Model

```{r}
print(mod)


rmodel = getLearnerModel(mod)
```

### Predict

```{r}
pred = predict(mod, task = task, subset = test)
head(as.data.frame(pred))


head(getPredictionProbabilities(pred))
```

### Performance

```{r}
performance(pred, measures = list(mlr::acc, mlr::auc))
```

### External Validation Set

Predict on data not included in the task

```{r}
test.data = dropNamed(data[test, ], "Survived")
pred = predict(mod, newdata = data[test, ])
performance(pred, measures = list(mlr::acc, mlr::auc))

```



